---
title: "Final Project Dynamic Regression"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Loading and Preparation

```{r include=FALSE}
# loading libraries
library(dplyr)
library(fpp2)
library(forecast)
library(tseries)
library(kader)
```

```{r}
# loading dataset
df <- read.csv('retail_sales.csv')
```

We need to get rid of 2020 since it is an outlier and would mess with our model since it is so different from other years.
```{r}
tail(df)
```

We need to get rid of 2020 since it is an outlier and would mess with our model since it is so different from other years.
```{r}
df <- df %>% filter(row_number() <= n()-8)
```

```{r}
tail(df)
```


```{r}
df <- df %>% select(alcohol_sales, sporting_goods_sales, clothing_sales, furniture_sales, restaurants_sales)
head(df)
```
# Clothing
```{r}
clothing <- ts(df$clothing_sales, frequency = 12, start = c(1992,01), end = c(2019,12))
autoplot(clothing)
```
```{r}
clothing_stl <- stl(clothing, s.window=12)
autoplot(clothing_stl) + 
  xlab("Year") + 
  ggtitle("Monthly Clothing Sales Decomposition")
```

# Sporting Goods
```{r}
sporting_goods <- ts(df$sporting_goods_sales, frequency = 12, start = c(1992,01), end = c(2019,12))
autoplot(sporting_goods)
```
```{r}
sporting_goods_stl <- stl(sporting_goods, s.window=12)
autoplot(sporting_goods_stl) + 
  xlab("Year") + 
  ggtitle("Monthly Sporting Goods Sales Decomposition")
```

# Furniture 
```{r}
furniture <- ts(df$furniture_sales, frequency = 12, start = c(1992,01), end = c(2019,12))
autoplot(furniture)
```
```{r}
furniture_stl <- stl(furniture, s.window=12)
autoplot(furniture_stl) + 
  xlab("Year") + 
  ggtitle("Monthly Furniture Sales Decomposition")
```

# Restaurants
```{r}
restaurant <- ts(df$restaurants_sales, frequency = 12, start = c(1992,01), end = c(2019,12))
autoplot(restaurant)
```
```{r}
restaurant_stl <- stl(restaurant, s.window=12)
autoplot(restaurant_stl) + 
  xlab("Year") + 
  ggtitle("Monthly Restaurants Sales Decomposition")
```

# Alcohol Sales
```{r}
alcohol <- ts(df$alcohol_sales, frequency = 12, start = c(1992,01), end = c(2019,12))
autoplot(alcohol)
```
```{r}
alcohol_stl <- stl(alcohol, s.window=12)
autoplot(alcohol_stl) + 
  xlab("Year") + 
  ggtitle("Monthly Alcohol Sales Decomposition")
```

# Variance Stabilizing and Sesonally adjusting
```{r}
lambda.opt <- BoxCox.lambda(clothing)
stable_clothing <- BoxCox(clothing, lambda=lambda.opt)
autoplot(stable_clothing)
lambda.opt
```
```{r}
stable_clothing_stl <- stl(stable_clothing, s.window=12)
```

```{r}
seasadj_clothing <- seasadj(stable_clothing_stl)
autoplot(seasadj_clothing) + ggtitle("Seasonally Adjusted Monthly Clothing Sales") + ylab("Sales (in millions of dollars)") + xlab('Year')
```
```{r}
lambda.opt <- BoxCox.lambda(furniture)
stable_furniture <- BoxCox(furniture, lambda=lambda.opt)

stable_furniture_stl <- stl(stable_furniture, s.window=12)

autoplot(stable_furniture)
lambda.opt
```
```{r}
seasadj_furniture <- seasadj(stable_furniture_stl)
autoplot(seasadj_furniture) + ggtitle("Seasonally Adjusted Monthly Furniture Sales") + ylab("Sales (in millions of dollars)") + xlab('Year')
```


```{r}
lambda.opt <- BoxCox.lambda(alcohol)
stable_alcohol <- BoxCox(alcohol, lambda=lambda.opt)

stable_alcohol_stl <- stl(stable_alcohol, s.window=12)

autoplot(stable_alcohol)
lambda.opt
```
```{r}
seasadj_alcohol <- seasadj(stable_alcohol_stl)
autoplot(seasadj_alcohol) + ggtitle("Seasonally Adjusted Monthly Alcohol Sales") + ylab("Sales (in millions of dollars)") + xlab('Year')
```
```{r}
lambda.opt <- BoxCox.lambda(sporting_goods)
stable_sporting_goods <- BoxCox(sporting_goods, lambda=lambda.opt)
lambda.opt
stable_sporting_stl <- stl(stable_sporting_goods, s.window=12)
```
```{r}
seasadj_sporting <- seasadj(stable_sporting_stl)
autoplot(seasadj_sporting) + ggtitle("Seasonally Adjusted Monthly Sporting Goods Sales") + ylab("Sales (in millions of dollars)") + xlab('Year')
```
```{r}
lambda.opt <- BoxCox.lambda(restaurant)
lambda.opt
stable_restaurants <- BoxCox(restaurant, lambda=lambda.opt)
stable_restaurants_stl <- stl(stable_restaurants, s.window=12)

autoplot(stable_)
```
```{r}
seasadj_restaurants <- seasadj(stable_restaurants_stl)
autoplot(seasadj_restaurants) + ggtitle("Seasonally Adjusted Monthly Restaurants Sales") + ylab("Sales (in millions of dollars)") + xlab('Year')
```



# Classic Linear Regression

```{r}
# combine the data
df <- cbind(seasadj_clothing, seasadj_alcohol, seasadj_furniture, seasadj_restaurants, seasadj_sporting)
head(df)
```

```{r}
lm.obj <- lm(seasadj_clothing ~ ., data=df)
summary(lm.obj)
```
```{r}
vif(lm.obj)
```
Restaurants sales is the highest VIF, therefore we need to drop it

```{r}
lm.obj <- lm(seasadj_clothing ~ seasadj_alcohol + seasadj_furniture + seasadj_sporting, data=df)
summary(lm.obj)
```
```{r}
vif(lm.obj)
```
After dropping restaurants, we still need to drop sporting goods.

```{r}
lm.obj <- lm(seasadj_clothing ~ seasadj_alcohol + seasadj_furniture, data=df)
summary(lm.obj)
```
```{r}
vif(lm.obj)
```
Now VIF values look good.

$$clothing_t = \beta_{0} + \beta_1 * alcohol_t  + \beta_2 * furniture_t + \epsilon_t $$
$$ \epsilon_t \sim iid N(0,\sigma^2)$$

```{r}
# Residual diagnostics
plot(lm.obj, which=1)
plot(lm.obj, which=2)
```
```{r}
df %>%
  as.data.frame %>%
  select(-seasadj_restaurants, -seasadj_sporting) %>%
  GGally::ggpairs()
```

```{r}
checkresiduals(lm.obj)
```

```{r}
tsdisplay(resid(lm.obj))
```
ACF is clearly slowly decreasing while PACF shows 3 spikes and quick drop offs. This is a clear sign that we must difference our errors
```{r}
hand.picked <- Arima(df[,"seasadj_clothing"],
                       xreg=df %>% as.data.frame %>% select(-seasadj_restaurants, -seasadj_sporting, -seasadj_clothing) %>%
                         as.matrix, order=c(0,1,0))
```


```{r}
checkresiduals(hand.picked)
```

```{r}
tsdisplay(diff(resid(lm.obj)), lag.max = 48)
```

```{r}
auto.fit <- auto.arima(df[,"seasadj_clothing"],
                       xreg=df %>% as.data.frame %>% select(-seasadj_restaurants, -seasadj_sporting, -seasadj_clothing) %>%
                         as.matrix)
auto.fit
```
```{r}
checkresiduals(auto.fit)
```
$$ clothing' =  \beta_1 * alcohol'_t + \beta_2 * furniture'_t $$
$$\epsilon_t  \sim ARIMA(1,1,1)(0,0,2)[12] errors $$
$$ \epsilon_t = \phi_1 \epsilon_{t-1} + \theta_1 \eta_{t-1} + \xi_{1} \eta_{t-12} + \xi_{2} \eta_{t-24} + \eta_t $$
$$ \eta_t \sim iid N(0,\sigma^2)$$

## Fitted equation:

$$ \hat{clothing}' =  43.38 * alcohol'_t + 2.02 * furniture'_t + \epsilon_t$$
$$\hat{\epsilon_t} =  $$

```{r}
coeftest(auto.fit)
```

```{r}
confint(auto.fit)
```

# After 2008 recession
```{r}
post_2008 <- window(df,start=c(2010,1), end=c(2019,12))
```

```{r}
autoplot(post_2008, facets = TRUE)
```


```{r}
lm.obj <- lm(seasadj_clothing ~ ., data=post_2008)
summary(lm.obj)
```
```{r}
vif(lm.obj)
```
```{r}
lm.obj <- lm(seasadj_clothing ~ seasadj_furniture + seasadj_restaurants + seasadj_sporting, data=post_2008)
summary(lm.obj)
```
```{r}
vif(lm.obj)
```
```{r}
lm.obj <- lm(seasadj_clothing ~  seasadj_restaurants + seasadj_sporting, data=post_2008)
summary(lm.obj)
```
```{r}
vif(lm.obj)
```

```{r}
checkresiduals(lm.obj)
```
```{r}
plot(lm.obj, which=1)
plot(lm.obj, which=2)
```


```{r}
tsdisplay(diff(resid(lm.obj)), lag.max=48)
```
```{r}
hand.picked <- Arima(post_2008[,"seasadj_clothing"],
                       xreg=post_2008 %>% as.data.frame %>% select(seasadj_restaurants, seasadj_sporting) %>%
                         as.matrix, order=c(0,1,0))

hand.picked
```


```{r}
auto.fit <- auto.arima(post_2008[,"seasadj_clothing"],
                       xreg=post_2008 %>% as.data.frame %>% select(seasadj_restaurants, seasadj_sporting) %>%
                         as.matrix)
auto.fit
```
```{r}
checkresiduals(auto.fit)
```
```{r}
coeftest(auto.fit)
```


$$ clothing_t =  \beta_0 + \beta_1 * restaurants_t + \beta_2 * sporting_t + \epsilon_t$$

$$\epsilon_t  \sim ARIMA(1,0,1) errors $$

$$ \epsilon_t = \phi_1 \epsilon_{t-1} + \theta_1 \eta_{t-1} + \eta_t $$

$$ \eta_t \sim iid N(0,\sigma^2)$$
```{r}
confint(auto.fit)
```

